Structure constants and short-time simulations of open collective quantum spin systems

### 2019-03-20 soc_squeezing/collective_paper/collective_spin.tex
added discussion of advantages and limitations of the truncated Taylor expansion method

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology,
quantum simulation, and quantum information.  Theoretical interest in
such systems dates back to the mid-twentieth century with the
introduction of the Dicke model\cite{dicke1954coherence} that
describes atoms cooperatively interacting with a single mode of the
radiation field and the Lipkin-Meshkov-Glick (LMG) model, a toy model
for testing many-body approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including atomic
ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
  schleier-smith2010states, chen2011conditional}, Bose-Einstein
condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
  klinder2015dynamical}, ultracold Fermi
gasses\cite{martin2013quantum, smale2018observation}, trapped
ions\cite{bohnet2016quantum}, optical
cavities\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement, hosten2016quantum,
  zhiqiang2017nonequilibrium, lewis-swan2018robust,
  norcia2018cavitymediated}, and molecular
sensors\cite{jones2009magnetic}.  These implementations have triggered
studies of a variety of rich subjects, including quantum
criticality\cite{latorre2005entanglement, alcalde2007functional,
  morrison2008dissipationdriven, morrison2008collective,
  sarandy2009classical, wang2012quantum, kessler2012dissipative,
  majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  bhattacherjee2014nonequilibrium, klinder2015dynamical,
  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
  lang2018concurrence, smale2018observation}, and precision
metrology\cite{wineland1992spin, itano1993quantum,
  kitagawa1993squeezed, ma2011quantum, takano2009spin,
  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
  zhong2010simplified, agarwal1997atomic, lau2014proposal,
  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
  schleier-smith2010squeezing, leroux2010implementation,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  jones2009magnetic}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
  ma2011quantum}.  Collective spin systems enable a means to break
through this limit via the preparation of many-body entangled states
such as NOON\cite{jones2009magnetic, chen2010heralded,
  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
  lau2014proposal, huang2015quantum}, and spin-squeezed
states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
  appel2009mesoscopic, schleier-smith2010states,
  schleier-smith2010squeezing, leroux2010implementation,
  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  ma2011quantum} that allow for measurement errors
$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
states typically requires nonlinear dynamics that are realized using
collisional, photon-mediated, or phonon-mediated interactions.
Although a truly collective spin model requires uniform, all-to-all
interactions, as long as measurements do not distinguish between
constituent particles even non-uniform systems can be effectively
described by a uniform model with renormalized
parameters\cite{hu2015entangled}.

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superseletion sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence will generally violate total spin conservation
and require the use of density operators, increasing the dimension of
accessible state space to $\sim N^3$\cite{hartmann2016generalized,
  xu2013simulating} and allowing for exact simulations of
$N\lesssim100$ particles.  If decoherence is sufficiently weak,
dynamics can be numerically solvable for $N\lesssim10^5$ particles via
``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
  zhang2018montecarlo} (also known as ``quantum jump'' or ``Monte
Carlo wavefunction'' methods) that can reproduce all expectation
values of interest.  When decoherence is strong, however, or when the
number of jump operators grows extensively with system size (e.g.~for
single-spin decay), these Monte Carlo methods can take a prohibitively
long time to converge, as simulations become dominated by incoherent
jumps that generate large numbers of distinct quantum trajectories
that need to be averaged to compute expectation values.

In this work, we develop a new method to simulate short-time dynamics
of collective spin systems with permutationally-symmetric decoherence.
This method is based on a short-time expansion of exact solutions to
the equations of motion for Heisenberg operators.  Evaluating this
expansion requires using the structure constants of a collective spin
operator algebra; the calculation of these structure constants is one
of the main technical results of this work, which we hope will enable
future studies to develop new analytical insights into the dynamics of
collective spin systems.

[Paragraph or so about benchmarking and the discussion section?]

@@

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the Appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin Hamiltonians in the form
\begin{align}
  H = \sum_{\v m} h_{\v m} \S_{\v m},
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{\v m}^*=h_{\v m^*}$ for
$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The time evolution of a particular
Heisenberg operator $\S_{\v m}$ under a Hamiltonian of the form in
Eq.~\eqref{eq:H_general} is then given by
\begin{align}
  \f{d}{dt}\S_{\v m}
  &= i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}} \\
  &= \sum_{\v n} T_{\v m\v n} \S_{\v n},
  \label{eq:EOM}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
with a corresponding decoherence rate $\gamma_\J$;
\begin{align}
  \D\p{\J;\O}
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
\end{align}
is the Heisenberg picture dissipator, or Lindblad superoperator, for
all jump operators $J\in\J$; and $T_{\v m\v n}$ is a matrix element of
the time derivative operator.  Decoherence via uncorrelated decay of
individual spins at a rate $\gamma_-$, for example, would be described
by the set of jump operators
$\J_-=\set{\sigma_-^{(j)}:j=1,2,\cdots,N}$ with
$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM} can be
determined from the structure constants $f_{\v n\v m\v\ell}$ of the
general product
$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}\S_{\v\ell}$ worked
out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
the effects of decoherence from jump operators (i.e.~elements of $\J$)
of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are respectively worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  These
ingredients are sufficient to compute matrix elements $T_{\v m\v n}$
of the time derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most
cases of experimental relevance.

@@

In principle, the factorial suppression of terms at higher orders
(i.e.~in $k$) of the expansion in Eq.~\eqref{eq:order_truncation}
implies for any given time $t$, there exists a truncation order $M_t$
for which truncation error is negligibly small.  In practice, only a
fixed maximal truncation order $M_{\t{max}}$ is accessible with
limited computational resources, such that
Eq.~\eqref{eq:order_truncation} only allows computation of the
correlator $\bk{\S_{\v m}}$ to a maximal time $t_{M_{\t{max}}}$.  The
maximal accessible truncation order $M_{\t{max}}$ generally depends on
the choice of operator basis for collective spin systems, which we
discuss briefly in Section \ref{sec:breakdown} and Appendix
\ref{sec:new_basis}.  As we will see in the following sections, our
choice of basis elements $\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$
allows us to accurately compute spin correlators to times that are
relevant e.g.~for spin squeezing protocols.

\section{Spin squeezing}

@@

The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
directionally-unbiased Ramsey squeezing parameter determined by the
maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
that achieved by any spin-polarized product state
(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
\begin{align}
  \xi^2
  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
  = \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $\v S\equiv\p{S_\z,S_\x,S_\y}$ is a collective spin vector, and
the minimization is performed over all unit vectors $\uv n$ in the
plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
parameter is entirely determined by one- and two-spin correlators of
the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
dynamics discussed here, these correlators are obtainable via exact
simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
projection $m$ onto the $z$ axis, i.e.~with
$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
permutationally-symmetric decoherence, these correlators are
obtainable with the collective-spin quantum trajectory Monte Carlo
method developed in ref.~\cite{zhang2018montecarlo}.  In this work,
these exact and quantum trajectory simulations will be used to
benchmark the expansion in Eq.~\eqref{eq:order_truncation}.

\begin{figure}
  \centering
  \subfloat[Squeezing under coherent dynamics \label{fig:coherent}]
  {\includegraphics{coherent.pdf}} \\
  \subfloat[Squeezing with decoherence:
  $\gamma_-=\gamma_+=\gamma_\z=\chi$ \label{fig:incoherent}]
  {\includegraphics{decoherence_weak.pdf}}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under \sref{fig:coherent} coherent and \sref{fig:incoherent}
    incoherent evolution, as computed using exact methods (solid
    lines), quantum trajectory simulations (dots), and the expansion
    in Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed
    lines).  Solid circles mark the times at which the last of these
    methods gives an unphysical result with $\xi^2<0$.}
  \label{fig:benchmarking}
\end{figure}

Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
both benchmarking simulations and the prescription in
Eq.~\eqref{eq:order_truncation} with $M=40$.  This squeezing is shown
for both coherent dynamics (Figure \ref{fig:coherent}), as well as
dynamics in the presence of uncorrelated emission, absorption, and
dephasing of individual spins at a rate $\chi$ (Figure
\ref{fig:incoherent}), respectively described by the sets of jump
operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence rates
$\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The squeezing results shown
in Figure \ref{fig:benchmarking} were computed in a rotated basis with
$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$; the only effect of this
rotation on the results presented in this figure is to roughly double
the time for which all TNT results agree.  The reason for different
results in rotated bases has to do with the breakdown of the expansion
in Eq.~\eqref{eq:order_truncation}, which we discuss further in
Section \ref{sec:breakdown}.

[Comment on near-perfect agreement right up until a sudden and drastic
departure from the correct result.  Reference figure showing squeezing
for large system with strong decoherence.  Mention that this method
was used to put a lower bound on squeezing via TAT in our
SOC+interactions squeezing paper.]

\section{Out-of-time-ordered correlators}

@@

\section{Breakdown}
\label{sec:breakdown}

@@

  Hermitian conjugation.  In any case, we express the final result of
  this section in both ordering conventions, and thus feel no need to
  reproduce these calculations with a different ordering.}.  This
choice of ordering compels us to expand

>>>>

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology and
quantum simulation.  Theoretical interest in such systems dates back
to the mid-twentieth century with the introduction of the Dicke
model\cite{dicke1954coherence} that describes atoms cooperatively
interacting with a single mode of a radiation field, and the
Lipkin-Meshkov-Glick (LMG) model, a toy model for testing many-body
approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including atomic
ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
  schleier-smith2010states, chen2011conditional}, Bose-Einstein
condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
  klinder2015dynamical}, ultracold Fermi
gasses\cite{martin2013quantum, smale2018observation}, trapped
ions\cite{bohnet2016quantum}, optical
cavities\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement, hosten2016quantum,
  zhiqiang2017nonequilibrium, lewis-swan2018robust,
  norcia2018cavitymediated}, and molecular
sensors\cite{jones2009magnetic}.  These implementations have triggered
studies of a variety of rich subjects, including quantum
criticality\cite{latorre2005entanglement, alcalde2007functional,
  morrison2008dissipationdriven, morrison2008collective,
  sarandy2009classical, wang2012quantum, kessler2012dissipative,
  majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  bhattacherjee2014nonequilibrium, klinder2015dynamical,
  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
  lang2018concurrence, smale2018observation}, and precision
metrology\cite{wineland1992spin, itano1993quantum,
  kitagawa1993squeezed, ma2011quantum, takano2009spin,
  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
  zhong2010simplified, agarwal1997atomic, lau2014proposal,
  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
  schleier-smith2010squeezing, leroux2010implementation,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  jones2009magnetic, he2019engineering}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
  ma2011quantum}.  Collective spin systems enable a means to break
through this limit via the preparation of many-body entangled states
such as NOON\cite{jones2009magnetic, chen2010heralded,
  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
  lau2014proposal, huang2015quantum}, and spin-squeezed
states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
  appel2009mesoscopic, schleier-smith2010states,
  schleier-smith2010squeezing, leroux2010implementation,
  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  ma2011quantum, he2019engineering} that allow for measurement errors
$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
states typically requires nonlinear dynamics that are realized using
collisional, photon-mediated, or phonon-mediated interactions.
Although a truly collective spin model requires uniform, all-to-all
interactions, as long as measurements do not distinguish between
constituent particles, even non-uniform systems can be effectively
described by a uniform model with renormalized
parameters\cite{hu2015entangled}.

In the absence of decoherence, permutation symmetry and total spin
conservation divide the total Hilbert space of a collective spin
system into superseletion sectors that grow only linearly with system
size $N$, thereby admitting efficient classical simulation of its
dynamics.  Decoherence will generally violate total spin conservation
and require the use of density operators, increasing the dimension of
accessible state space to $\sim N^3$\cite{hartmann2016generalized,
  xu2013simulating} and allowing for exact simulations of
$N\lesssim100$ particles.  If decoherence is sufficiently weak,
dynamics can be numerically solvable for $N\lesssim10^5$ particles via
``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
  zhang2018montecarlo} (also known as ``quantum jump'' or ``Monte
Carlo wavefunction'' methods) that can reproduce all expectation
values of interest.  When decoherence is strong, however, or when the
number of jump operators grows extensively with system size (e.g.~for
single-spin decay), these Monte Carlo methods can take a prohibitively
long time to converge, as simulations become dominated by incoherent
jumps that generate large numbers of distinct quantum trajectories
that need to be averaged over to compute expectation values.

In this work, we present a new method to simulate short-time dynamics
of collective spin systems with permutationally-symmetric decoherence.
This method is based on a short-time expansion of exact solutions to
the equations of motion for Heisenberg operators.  Evaluating this
expansion requires using the structure constants of a collective spin
operator algebra; the calculation of these structure constants (in
Appendices \ref{sec:identities}--\ref{sec:prod_general}) is one of the
main technical results of this work, which we hope will enable future
studies to develop new analytical insights into the behavior of
collective spin systems.  We benchmark our method against both exact
and quantum trajectory Monte Carlo calculations of spin squeezing in
accessible parameter regimes, highlighting both advantages and
limitations of the short-time expansion.  Finally, we showcase an
application of our method with calculations that are inaccessible to
other methods.

[More text about the content of this paper.]

@@

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin Hamiltonians in the form
\begin{align}
  H = \sum_{\v m} h_{\v m} \S_{\v m},
  \label{eq:H_general}
\end{align}
where Hermiticity requires that $h_{\v m}^*=h_{\v m^*}$ for
$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The time evolution of a particular
Heisenberg operator $\S_{\v m}$ under a Hamiltonian of the form in
Eq.~\eqref{eq:H_general} is then given by
\begin{align}
  \f{d}{dt}\S_{\v m}
  &= i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}} \\
  &= \sum_{\v n} T_{\v m\v n} \S_{\v n},
  \label{eq:EOM}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
with a corresponding decoherence rate $\gamma_\J$;
\begin{align}
  \D\p{\J;\O}
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
\end{align}
is the Heisenberg picture dissipator, or Lindblad superoperator, for
all jump operators $J\in\J$; and $T_{\v m\v n}$ is a matrix element of
the time derivative operator.  Decoherence via uncorrelated decay of
individual spins at a rate $\gamma_-$, for example, would be described
by the set of jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$
with $\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM}
can be determined from the structure constants $f_{\v\ell\v m\v n}$ of
the general product
$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ worked
out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
the effects of decoherence from jump operators (i.e.~elements of $\J$)
of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
consider these calculations one of the main technical contributions of
our work, with potential applications beyond the short-time simulation
method presented here.  These ingredients are sufficient to compute
matrix elements $T_{\v m\v n}$ of the time derivative operator $d/dt$
in Eq.~\eqref{eq:EOM} in most cases of practical interest.

@@

In principle, the factorial suppression of terms at higher orders of
the Taylor expansion in Eq.~\eqref{eq:time_series} implies for any
given time $t$, there exists a truncation order $M_t$ for which the
truncation error in Eq.~\eqref{eq:order_truncation} is negligibly
small.  In practice, only a maximal truncation order $M_{\t{max}}$ is
accessible with limited computational resources, such that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
allows computation of the correlator $\bk{\S_{\v m}}$ to a maximal
time $t_{M_{\t{max}}}$.  As we will see in the following section,
$M_{\t{max}}=40$ will suffice for the accurate computation of
collective spin correlators up to times that are relevant e.g.~for
spin squeezing protocols.

\section{Spin squeezing, benchmarking, and breakdown}

@@

The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
directionally-unbiased Ramsey squeezing parameter determined by the
maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
that achieved by any spin-polarized product state
(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
\begin{align}
  \xi^2
  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
  = \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $\v S\equiv\p{S_\x,S_\y,S_\z}$ is a collective spin vector, and
the minimization is performed over all unit vectors $\uv n$ in the
plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
parameter is entirely determined by one- and two-spin correlators of
the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
dynamics discussed here, these correlators are obtainable via exact
simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
projection $m$ onto the $z$ axis, i.e.~with
$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
permutationally-symmetric decoherence, these correlators are
obtainable with the collective-spin quantum trajectory Monte Carlo
method developed in ref.~\cite{zhang2018montecarlo}.  In this work,
these exact and quantum trajectory simulations will be used to
benchmark the truncated Taylor expansion in
Eq.~\eqref{eq:order_truncation}.

\begin{figure}
  \centering
  \subfloat[Squeezing under coherent dynamics
  \label{fig:squeezing_coherent}]
  {\includegraphics{coherent.pdf}} \\
  \subfloat[Squeezing with decoherence:
  $\gamma_-=\gamma_+=\gamma_\z=\chi$
  \label{fig:squeezing_incoherent}]
  {\includegraphics{decoherence_weak.pdf}}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under \sref{fig:squeezing_coherent} coherent and
    \sref{fig:squeezing_incoherent} incoherent evolution, computed
    using exact methods (solid lines), quantum trajectory simulations
    (dots), and the truncated Taylor expansion in
    Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed lines).
    Solid circles mark the times at which the truncated Taylor
    expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:benchmarking}
\end{figure}

Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
both benchmarking simulations and the truncated Taylor expansion in
Eq.~\eqref{eq:order_truncation} with $M=40$.  Squeezing is shown for
both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
as dynamics in the presence of uncorrelated emission, absorption, and
dephasing of individual spins at a rate $\chi$ (Figure
\ref{fig:squeezing_incoherent}), respectively described by the sets of
jump operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence
rates $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The results shown in
Figure \ref{fig:benchmarking} were computed in a rotated basis with
$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
appropriate transformations of the Hamiltonian and jump operators);
the only effect of this rotation on the results presented in Figure
\ref{fig:benchmarking} is to prolong the time for which the truncated
Taylor expansion agrees with the benchmarking simulations.  The reason
for different results in rotated basis has to do with the breakdown of
the truncated Taylor expansion, which we discuss further below.

The main lesson from Figure \ref{fig:benchmarking} is that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} gives
essentially exact results right up until a sudden and drastic
departure that is simple to diagnose by inspection.  This drastic
departure occurs because the truncated Taylor expansion neglects
high-weight (i.e.~large-$\abs{\v n}$) operators $\S_{\v n}$ whose
contributions to a correlator of interest $\bk{\S_{\v m}}$ eventually
become non-negligible.  For sufficiently large times $t$, these
individual contributions may generally be large compared to the value
of $\bk{\S_{\v m}}$, in which case their truncation will yield
non-physical results.  This breakdown mechanism is also the reason for
prolonged agreement between the truncated Taylor expansion and
benchmarking methods when squeezing the initial state $\ket{-\Z}$
rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
correlators $\bk{\S_{\v n}}_{t=0}$ vanish unless $n_+=n_-=0$ (see
Appendix \ref{sec:initial_conditions}), so there is a substantially
smaller number of terms with non-zero contribution to the correlator
$\bk{\S_{\v m}}$, and therefore a substantially smaller number of
non-zero terms that are neglected in the truncated Taylor expansion.

Although the truncated Taylor expansion breaks down at short times, it
has two key advantages over other methods to compute collective spin
correlators.  First, computing spin correlators with the truncated
Taylor expansion is generally much faster and requires much fewer
computing resources than alternatives.  The quantum trajectory Monte
Carlo simulations performed for Figure \ref{fig:squeezing_incoherent},
for example, take $\sim10^4$ CPU hours to compute on standard modern
computing hardware; the bulk of this time is spent performing sparse
matrix multiplication, leaving little room to further optimize run
time.  Parallelization can reduce the actual run time of these quantum
trajectory simulations to $\sim10$ hours, but at the cost of greatly
increasing computing resource requirements.  The truncated Taylor
expansion results in Figure \ref{fig:squeezing_incoherent}, meanwhile,
take $\sim10$ seconds to compute with a single CPU on similar
computing hardware.

Second, the truncated Taylor expansion enables computing spin
correlators in strong-decoherence regimes entirely inaccessible to
other methods.  As an example, Figure \ref{fig:decoherence_strong}
shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, with decay
and dephasing of individual spins at rates
$\gamma_-=\gamma_\z=100\chi$.  Computations of this sort were used to
put lower bounds on theoretically achievable spin squeezing via TAT
with decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
concrete application of the truncated Taylor expansion and the
collective-spin structure constants calculated in this work.

\begin{figure}
  \centering
  \includegraphics{decoherence_strong.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under incoherent evolution with decay and dephasing of individual
    spins at rates $\gamma_-=\gamma_\z=100\chi$, computed using the
    truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} with
    $M=40$.  Solid circles mark the times at which the truncated
    Taylor expansion gives an unphysical result with $\xi^2<0$.}
  \label{fig:decoherence_strong}
\end{figure}

\section{Out-of-time-ordered correlators}
\label{sec:OTOC}

@@

@@

  Hermitian conjugation.  In any case, we provide the final result of
  this section in both ordering conventions, and therefore feel no
  need to reproduce these calculations with a different ordering of
  spin operators.}.  This choice of ordering compels us to expand
